### 2022-01-16 newvariant.tex
more checks

Recent studies have suggested the possibility that new variants may have different generation-interval distributions.
For example, Kissler \textit{et al.} \citep{kissler2021densely} suggested that the Alpha variant may have a longer duration of infection: 13.3 days (90\% CI: 10.1--16.5), compared to 8.2 days (90\% CI: 6.5--9.7) for the wild type, thus suggesting that the mean generation interval of the Alpha variant is likely to be longer than that of the wild type \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model}.
In contrast, some studies have suggested that the Delta variant may have a shorter generation interval due to faster within-host viral replication \citep{li2021viral,zhang2021transmission}.
Modeling studies have also considered the possibility that the fast replacement seen in some variants may be driven, in part, by shorter generation intervals \citep{davies2021estimated,volz2021transmission}.
However, linking strength and speed is complicated given that the generation-interval distribution depends on many factors including behavior:
for example, self-isolation after symptom onset will lead to shorter generation intervals.
The recent emergence of the variant, and its breakthrough infections in previously immune individuals, adds further uncertainty to individual-level transmission dynamics and therefore the generation-interval of SARS-CoV-2 variants \citep{pulliam2021increased}.

@@

We explored how the generation-interval distribution shapes the link between relative strength and speed.
While the role of generation-interval distributions in linking epidemic strength and speed has been well-established, this framework can also provide provides insight into an ut under-appreciated concept: the relative strength and speed of new variants also depend on both current epidemic conditions and the nature of control measures (e.g., strength- or speed-like) to be introduced.
For example, under a constant-strength intervention that reduces the transmission rate of the wild type and a variant by the same amount (and therefore, keeping the relative strength constant), the relative speed of the variant is sensitive to the epidemic growth rate of the wild type strain.
Our results challenge the commonly made assumption that relative speed of new variants remains constant over time when characterizing changes in the proportion of new variants.

>>>>

Recent studies have suggested the possibility that new variants may have different generation-interval distributions.
For example, Kissler \textit{et al.} \citep{kissler2021densely} suggested that the Alpha variant may have a longer duration of infection: 13.3 days (90\% CI: 10.1--16.5), compared to 8.2 days (90\% CI: 6.5--9.7) for the wild type, thus suggesting that the mean generation interval of the Alpha variant is likely to be longer than that of the wild type \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model}.
In contrast, some studies have suggested that the Delta variant may have a shorter generation interval due to faster within-host viral replication \citep{li2021viral,zhang2021transmission}.
Modeling studies have also considered the possibility that the fast replacement seen in some variants may be driven, in part, by shorter generation intervals \citep{davies2021estimated,volz2021transmission}.
However, linking strength and speed is complicated given that the generation-interval distribution depends on many factors including behavior:
for example, self-isolation after symptom onset will lead to shorter generation intervals.
The recent emergence of the Omicron variant, and its breakthrough infections in previously immune individuals, adds further uncertainty to individual-level transmission dynamics and therefore the generation-interval of SARS-CoV-2 variants \citep{pulliam2021increased}.

@@

We explored how the generation-interval distribution shapes the link between relative strength and speed.
While the role of generation-interval distributions in linking epidemic strength and speed has been well-established, this framework can also provide provides insight into an under-appreciated concept: the relative strength and speed of new variants also depend on both current epidemic conditions and the nature of control measures (e.g., strength- or speed-like) to be introduced.
For example, under a constant-strength intervention that reduces the transmission rate of the wild type and a variant by the same amount (and therefore, keeping the relative strength constant), the relative speed of the variant is sensitive to the epidemic growth rate of the wild type strain.
Our results challenge the commonly made assumption that relative speed of new variants remains constant over time when characterizing changes in the proportion of new variants.
